---
title: "Z Table"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
**Quesetion-1**
Use R to build the positive half of the paper z table (from 0 to 3.99). Build each column separately. Then use cbind() or data.frame() to combine the columns into a single matrix or data.frame. Call it mymat, and name the rows and columns like in the paper table. At the end, type mymat to display your matrix. 


```{r}

#Answer(1)----------------------------------------

#build row and column labels.
myrows <- .1*(0:39)
mycols <- .01*(0:9)

#build the first column by passing its input vec to pnorm ().
vec <- myrows
mymatp1=pnorm(vec)

#build next columns with loop and combine columns into a matrix.
for(i in 1:9) {
  
    #make vec_next (the input of next column).
    vec_nex=vec+0.01*i
    
    #pass vec_nex to pnorm( ) and compute table column and combine it with last columns.
    mymatp1=cbind(mymatp1,pnorm(vec_nex))
}

#name rows and columns like the paper table.
rownames(mymatp1) <- myrows
colnames(mymatp1) <- mycols

#display mymat.
View(mymatp1) 

```

**Quesetion-2**
Construct mymat again, but this time build each row separately and use rbind() to append it to the table. There are many rows, so use a for{} loop. 

```{r}
#Answer(2)----------------------------------------

#build row and column labels.
myrows <-  seq(0,3.99, by=.1)
mycols <-  seq(0,0.09, by=.01)

#build the first row by passing its input vec to pnorm ().
vec2=mycols
mymatp2=pnorm(vec2)

#build next columns with loop and combine columns into a matrix.
for(i in 1:39) {
  
    #make vec_next (the input of next row).
    vec_nex=vec2+0.1*i
    
    #pass vec_nex to pnorm( ) and compute table row and combine it with last rows.
    mymatp2=rbind(mymatp2,pnorm(vec_nex))
    }

#name rows and columns like the paper table.
rownames(mymatp2) <- myrows
colnames(mymatp2) <- mycols

#display mymat.
View(mymatp2) 

```
**Quesetion-3**
One more time, but now ???ll the table one cell at a time. This approach will require two nested for{}loops. 

```{r}
#Answer(3)----------------------------------------

#build row and column labels.
myrows <- .1*c(0:39)
mycols <- .01*c(0:9)

#create the full-size matrix and fill its cells with zero.
mymatp3 <- matrix(0,nrow=40, ncol=10) 

#lable our matrix
rownames(mymatp3) <- myrows 	
colnames(mymatp3) <- mycols

# "i" will be rows' index, "j" will be columns' index.
for(i in 1:40){
  for(j in 1:10){
    
    #fill cell[i,j] of matrix with pnorm().
    mymatp3[i,j] <- pnorm(myrows[i]+mycols[j])
  }
} 

#display mymat.
View(mymatp3) 

```
**Quesetion-4**
No loops this time.

```{r}
#Answer(4)----------------------------------------

#create a vector with elements 0,.01,...,3.99.
xvec<- seq(0,3.99,0.01)

#build another vector containing pnorm() of the elements.
pvec<- pnorm(xvec)

#pass the vectors to the matrix function.
mymatp4=matrix(pvec,40,10,byrow = TRUE)

#build row and column labels.
myrows <- .1*(0:39)
mycols <- .01*(0:9)

#lable our matrix.
dimnames(mymatp4) <- list(myrows[1:40], mycols[1:10])

#display mymat.
View(mymatp4)

```
 
**Quesetion-5** 
Finally, we will build all of the paper table, both the negative and positive parts.


```{r}
#part(5)----------------------------------------

#create a vector with both negative and positive parts.
#First way
x<- seq(-3.99,0,0.01)
y<- seq(0,3.99,0.01)
xvec=c(x,y)

#Second way
xvec_1<-0.01*array(-399:399)
xvec_1=append(xvec_1,0,after=400)

#Third way
xvec_2<- 0.01*c(-399:400)
xvec_2[401:800]=xvec_2[401:800]-0.01


#build another vector containing pnorm() of the elements.
pvec<- round(pnorm(xvec),5)


#Negative part
#separate the probability of ngative part in our vecctor.
pvec_N=pvec[1:400]

#pass the vectors to the matrix function.
mymat_num=matrix(pvec_N,40,10,byrow = TRUE)

#unlike the positive part,in the negative part of paper table the probabilities 
#of each row are in decreasing order, so I will sort each part large to small.
#I multiply the entire matrix by negative one so that I can sort them from biggest 
#to smallest using apply_sort function.I transpose the matrix back to the original
#shape.
mymat_num_neg=-mymat_num
mymat5_1=-t(apply(mymat_num_neg,1,sort))

#build row and column labels.
mycols <- .01*(0:9)
myrows <- .1*(-39:0)

#lable our matrix.
dimnames(mymat5_1) <- list(myrows,mycols)


#positive part
#separate the probabilities of positive part in our vecctor.
pvec_p=pvec[401:800]

#pass the data to the matrix function.
mymat5_2=matrix(pvec_p,40,10,byrow = TRUE)

#build row and column labels.
myrows <- .1*(0:39)
mycols <- .01*(0:9)

#lable our matrix, just the row part.
dimnames(mymat5_2) <- list(myrows, NULL)


#rbind negative and positive rows together
mymat=rbind(mymat5_1,mymat5_2)

#display mymat.
mymat
```

