---
title: "Dice"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

**Question**

Consider rolling two fair dice; suppose one die is red and the other is blue. Let

**D1 = the outcome of the red die**
  
**D2 = the outcome of the blue die**
  
**Y = D1 + D2** 

Concise instructions: Use R to build the pmf (probability mass function) for D1 as a two-column data.frame, and use this table to compute its mean, variance (twice! see details below), and standard deviation. Repeat these steps for D2 and for Y . Use matrix multiplication wherever possible.



```{r}
#1. Build the pmf for D1------------------------------------------------------------

#possible outcomes.
k=c(1,2,3,4,5,6)

#probability of outcomes.
pk=c(1/6,1/6,1/6,1/6,1/6,1/6)

#build a data.frame.
pmfD1=data.frame(k,pk)
pmfD1

#verify the probabilities add up to one
sum(pmfD1$pk)

#2. Compute the mean----------------------------------------------------------------
mu1=k%*%pk
mu1

#3. compute variance using the defination------------------------------------------- 
var1_1=((k-as.vector(mu1))^2)%*%pk
var1_1

#using the shortcut formula of variance
var1_2=((k^2)%*%pk)-(mu1^2)
var1_2

#4. compute the standard deviation--------------------------------------------------
sd1=sqrt(var1_1)
sd1

#5. Repeat steps 1-4 for D2---------------------------------------------------------

k=c(1,2,3,4,5,6)
pk=c(1/6,1/6,1/6,1/6,1/6,1/6)

pmfD2=data.frame(k,pk)
pmfD2

sum(pmfD2$pk)

mu2=k%*%pk
mu2

var2_1=((k-as.vector(mu2))^2)%*%pk
var2_1

var2_2=((k^2)%*%pk)-(mu2^2)
var2_2

sd2=sqrt(var2_2)
sd2

#5. Repeat steps 1-4 for Y----------------------------------------------------------

#build row and column labels.
myrows <-c(1:6)
mycols <-c(1:6)

#create the full-size matrix and fill its cells with zero.
y <- matrix(0,nrow=6, ncol=6) 

#lable our matrix
rownames(y) <- myrows 	
colnames(y) <- mycols

# "i" will be rows' index, "j" will be columns' index.
for(i in 1:6){
  for(j in 1:6){
    
    #fill cell[i,j] of matrix.
    y[i,j] <-myrows[i]+mycols[j]
    

  }
} 

#display y.
View(y) 

#possible outcomes.
k=c(2:12)

#probability of outcomes:

##build a vector with 12 element of zero.
pk=0*c(1:11)

##fill the vector with probabilities.
for(r in 2:12)
{
  #count the number of occurrence.
  pk[r-1]=(sum(match(y,r,nomatch =0))/36)
}
pk

#build a data.frame.
pmfy=data.frame(k,pk)
pmfy  

#verify the probabilities add up to one  
sum(pmfy$pk)

#Compute the mean
mu=pk%*%k
mu

#compute variance using the defination
var_1=((k-as.vector(mu))^2)%*%pk
var_1

#using the shortcut formula of variance
var_2=((k^2)%*%pk)-(mu^2)
var_2

#compute the standard deviation
sd2=sqrt(var_2)
sd2



```







